\documentclass{article}

\RequirePackage{hyperref}

\usepackage[parfill]{parskip}
\usepackage[letterpaper, portrait, margin=1in]{geometry}
% Author
\usepackage{authblk}
% Images
\usepackage{graphicx}
\usepackage{lscape}
\usepackage{longtable}
% \usepackage{svg}

%\articlesubtype{This is the article type (optional)}
% \bibliography{paper-webserver.bib}

\usepackage{xcolor}
\newcommand\mwcomment[1]{\textcolor{red}{#1}}

\begin{document}

\title{K-mer based prokaryote pan-genome graph creation for long-term surveillance}

\author[1]{Kevin K Le\thanks{kle009@uottawa.ca}}
\author[1]{Victor PJ Gannon}
\author[2]{Chad R Laing\thanks{chad.laing@canada.ca}}
\affil[1]{National Microbiology Laboratory at Lethbridge, Public Health Agency of Canada, Twp Rd 9-1, Lethbridge, AB, T1J 3Z4, Canada}
\affil[2]{National Centre for Animal Diseases, Lethbridge Laboratory, Canadian Food Inspection Agency, Twp Rd 9-1, Lethbridge, AB, T1J 3Z4, Canada}

\renewcommand\Authands{ and }

\maketitle

\begin{abstract}

Current methods for pan-genome creation, which offer a callable mapping of all genes within a sample population, are based on the prior identification of known genes and produce one-off results which can not be updated as new samples are sequenced.
Compactation steps, which reduce the final output size and are common in pan-genome creation software, prevent updating the pangome after creation and individual whole-genome sequence samples from being reconstructed from a computed pan-genome.
Instead, an optimized graph structure could support traditional pan-genome queries even with a large output size.
Here we present a method for pan-genome creation based on an uncompressed De-Bruijn graph with edge metadata that functions similarly to common pan-genomes, supports updating the pan-genome in linear time, and returns source attribution during variant calling.


\end{abstract}


\section{Introduction}

% Paragraph on pan-genomes
Pan-genomes, the representation of all genes within a population, are important for distinguishing genetic components to observed phenotypes.
Traditional phylogenetic trees used to distinguish different subtypes by branch points may be poor genetic determinants in some species \cite{vernikos2015ten} because of the high degree of genetic exchange \cite{medini2005microbial}.
Instead, pan-genomes model a clade as based off \textit{core} conserved regions along with variable \textit{accesory} regions.

% Paragraph on graph structures
% text...

% Paragraph on need to constantly update
As new samples are sequenced however, the number of variants in accesory genes are expected to rise even if the core genes remain unchanged \cite{medini2005microbial}.
This is particularly relevant for some species such as \textit{Escherichia coli.} where the core genome only represents a small percentage of the overall genome \cite{fukiya2004extensive}.
Futhermore, for public health efforts and outbreak surveillance, pan-genomes data structures must be updatable in a reasonable timeframe.
The update requirement precludes a full re-computation of the pan-genome as such an approach would not scale to an ever growing dataset of genome samples.

% Paragraph for lead-in
% text...

% **************************************************************
% Keep this command to avoid text of first page running into the
% first page footnotes
\enlargethispage{-65.1pt}
% **************************************************************

\section{Related Work}

% Alginment based
% Panseq
% ???
The main focus of pan-genomics software is on the identification of core and accessory gene regions, with improvements trying to increase possible sample size or accomadating greater variation and differing reading frames.
Software such as Panseq \cite{laing2010pan} are based on alignment algorithmns, with core regions identified as coding genes with pair-wise sequence similarity above a certain cut-off.
Subsequent packages such as PanOCT \cite{fouts2012panoct} and GET_HOMOLOGUES \cite{contreras2013get_homologues} use alignment approaches for accepting sequence similarity along with clustering approaches for core region determination.

% Annotation based
% PGAP ???
% Roary
Later software packages integrated functional annotation either post-graph creation \cite{zhao2011pgap} or as an integral part of graph creation \cite{page2015roary}.
The Pan-Genomes Analysis Pipeline (PGAP) \cite{zhao2011pgap} reported pan-genome results for 11 samples, and an 

% Data structure
% SplitMEM

% Key problems and our approach
In all these cases the creation of a pan-genome follows a combinatorial increase in the number of comparisons with a larger sample size \cite{vernikos2015ten}, though the rate of new ortholog discovery decreases under Heap's law \cite{tettelin2008comparative}.
% Coding vs non-coding
The current spate of pan-genomics software have focused exclusively on coding regions of DNA.
While useful for reducing the size of input data for comparisons, evidence \cite{} has pointed to the contribution of non-coding regions to species phenotypes.


\section{Implementation}

\section{Results}

\section{Discussion}


\section{Conclusions}


\textit{Conflict of interest}. None declared.

\newpage

\bibliographystyle{unsrt}
\bibliography{paper}

\end{document}
